Evaluating approaches to deriving photometric redshift posterior probability distribution functions for \textsc{LSST}

### 2019-03-10 paper/tex/methods.tex
working through BPZ section

\section{Methods}
\label{sec:pzcodes}
%(Sam Schmidt, Rongpu Zhou, Ibrahim Almosallam, Eric Nuss, Johann Cohen Tanugi, John Soo)

Here we outline the photo-$z$ PDF codes tested in this study. In total, twelve distinct codes are tested.  This sample is not comprehensive, codes were chosen based on the expertise available within the group; however, those chosen do cover a broad range of the current-generation methods used in the field.  Both template-based and machine learning approaches are included and each are described separately in Secs.~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively. The list of codes are summarized in Table.~\ref{tab:list_of_codes}.  All code runners were asked to output redshift posterior estimates on 200 linear-spaced bins between redshifts 0 and 2.

\begin{table*}  %%% DATA TABLE %%%
\caption{List of photo-$z$ codes featured in this study. ML here means machine learning.} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
\begin{tabular}{llll}
\hline
\bf Code 			& \bf Type & \bf Paper & \bf Website \\
\hline
\textsc{BPz} 		& template 			& \citet{Benitez:00}		& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
\textsc{EAZY} 		& template 			& \citet{Brammer:08}		& \url{https://github.com/gbrammer/eazy-photoz} \\
\textsc{LePhare} 	& template 			& \citet{Arnouts:99}		& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
\textsc{ANNz2} 		& ML 	& \citet{Sadeh:16}			& \url{https://github.com/IftachSadeh/ANNZ} \\
\textsc{Delight} 	& ML/template 	& \citet{Leistedt:17}		& \url{https://github.com/ixkael/Delight} \\
\multirow{ 2}{*}{\textsc{FlexZBoost}} & \multirow{ 2}{*}{ML} 	& \multirow{ 2}{*}{\citet{Izbicki:17}}		& \url{https://github.com/tpospisi/flexcode};\\
					& 		&							& \url{https://github.com/rizbicki/FlexCoDE}\\
%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
\textsc{GPz} 		& ML	& \citet{Almosallam:15b}	& \url{https://github.com/OxfordML/GPz} \\
\textsc{METAPhoR} 	& ML 	& \citet{Cavuoti:17}		& \url{http://dame.dsf.unina.it}\\
\textsc{CMNN} 		& ML 	& \citet{Graham:17}			& - \\
\textsc{SkyNet} 	& ML 	& \citet{Graff:14}			& \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
\textsc{TPZ} 		& ML 	& \citet{Carrasco_Kind:13}	& \url{https://github.com/mgckind/MLZ} \\
\hline
\trainz 	& N/A	& See Section~\ref{sec:method:trainz} & \\
\end{tabular}}
\end{table*}

%\red{The questions that must be answered for each code are\dots}
The questions that must be answered for each code are: what unique features are included in the specific implementation that influence the output $p(z)$.  What form of validation was performed with the training data, how were photometric uncertainties employed in the analysis, how were negative fluxes treated, what specific prior form was employed (for template based codes), or what specific machine learning architecture was used (for ML codes)?
%\scc{which parameter (in a general meaning) have been used, which kind of validation has been performed (if any)}
\subsection{Template-based Approaches}
\label{sec:templatecodes}
We test three publicly available and commonly used template-based codes: \textsc{BPZ}, \textsc{EAZY}, and \textsc{LePhare}.  All three codes follow the standard procedure for template-based redshift estimation: calculate model fluxes for a set of template spectral energy distributions (SEDs) on a grid of redshift values and calculating a $\chi^2$ merit function using the observed and model fluxes:
\begin{equation} \label{eq_temp_chi}
\chi^2(z,T,A) = \sum_i^{N_{\rm filt}}\left(\frac{F^i_{\rm{obs}}−A \times F^i_{\rm{pred}}(T,z)}{\sigma^i_{\rm{obs}}}\right)^2
\end{equation}
\noindent  where $A$ is a normalization factor, $F^i_{\rm{pred}}(T,z)$ is the flux predicted for a template $T$ at redshift $z$. $F^i_{\rm{obs}}$ is the observed flux in a given band $i$ and $\sigma^i_{\rm{obs}}$ is the observed flux error. $N_{\rm tot}$ is the total number of filters, in our case the six $ugrizy$ LSST filters.  Specific implementation details of each code, e.~g.~prior form and implementation, are described below.

@@

\textsc{BPZ}\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based photo-$z$ code that compares the expected colors ($C$) calculated for a set of spectral energy distribution (SED) types/templates ($T$) to the observed colors to calculate the likelihood of observing colors at each redshift for each type, $p(C|z,T)$.  The likelihoods at each redshift are related to the $\chi^2$ in Equation~\ref{eq_temp_chi} by the simple form: likelihood $\propto \, e^{-\chi^2/2}$. The code employs an empirically determined Bayesian prior in apparent magnitude ($m_0$) and SED-type. Assuming that the SED-types are spanning and exclusive, we can determine the redshift posterior $p(z|C,m_0)$ by marginalizing over all SED-types with a simple sum \citep[Eq.~3 from][]{Benitez:00}:

\begin{equation} \label{eq:redshift_posterior}
p(z|C,m_0)\,\propto\, \sum_{T}p(z,T|m_0)\,p(C|z,T)
\end{equation}
\noindent where the first term on the right-hand side is the Bayesian prior and the second term is the traditional likelihood. The prior is assumed to have the form: $p(z,T|m_0)=p(T|m_0)\,p(z|T,m_0)$, i.e. it parameterizes the prior as an evolving type fraction with apparent magnitude, combined with a prior on the expected redshift probability distribution as a function of both apparent magnitude and SED-type.

In this paper we use \textsc{BPZ v 1.99.3}. The template set employed here is the set of $100$ discrete SEDs described in Section~\ref{sec:buzztemplates}
%, and the $129$ Brown SEDs for the Galacticus dataset.
To keep the number of free parameters to a manageable level the SEDs in the training set are sorted by the rest-frame $u$-$g$ colour and split into three ``broad'' SED classes, equivalent to the \texttt{E}, \texttt{Sp} and \texttt{Im/SB} types in \citet{Benitez:00}. We assume the same functional form for the Bayesian priors as used by \citet{Benitez:00}, and utilize the training-set galaxies with known SED-type, redshift, and apparent magnitude to determine the type fractions and the best fit for the eleven free parameters of the prior.
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
For galaxies that are not detected in a measured band, the placeholder value is replaced with an estimate of the one $\sigma$ detection limit in that particular band, i.~e.~a value close to the estimated sky noise threshold.
The type-marginalized $p(z)$ is generated by setting the parameter \texttt{PROBS\_LITE=TRUE} in the \textsc{BPZ} parameter file.

@@

In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates. We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry. %\red{The quality parameter \texttt{Qz} is...}
The likelihoods include the application of a type-independent apparent magnitude prior estimated from the training data.

@@

\subsection{Training-based Codes}
\label{sec:trainingcodes}
The training-based codes use a variety of algorithms in order to estimate $p(z)$, specifics of each implementation are described in the subsections.  Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.  Another key difference is the treatment of non-detections in one or more bands.  Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.  There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.  The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.  However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.

@@

\begin{equation}
p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
\end{equation}
\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$. For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:

\begin{equation} \label{eq:delight_noisy}
p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
\end{equation}
where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$. Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift. The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,

@@

\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.} The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:

\begin{equation}
f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
\end{equation}
By the orthogonality property, the expansion coefficients are just conditional means

\begin{equation}
\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
\end{equation}
These coefficients can easily be estimated from data by regression since $\mathbb{E}\left[\phi_i(z)|\x\right]$ is the regression of $\phi_i(z)$ (a transformation of $Z$) on $X$.

@@

The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing any missing magnitude values using a simple linear model that estimates the missing magnitudes given the observed ones. The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained. The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[

\begin{equation}
x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
\end{equation}  %\]

where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$). This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e. $\Sigma_{uo}=0$, this will reduce to a simple average predictor.  We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.

@@

\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.  The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.  The photometric perturbation is defined as:
\begin{equation}
 m_{ij} = m_{ij} + \alpha_{i}F_{ij}u_{\mu=0,\sigma=1}
\end{equation}
\noindent  where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

@@

\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}
In addition to the main photo-$z$ algorithms described above we also include a very simple method as a pathological example.  For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:

\begin{equation}
p(z) = \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}}z_{\mathrm train}
\end{equation}
That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.  This estimator is essentially a k nearest-neighbour estimator with k equal to the number of galaxies in the training sample.  As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical, modulo fluctuations due to finite sample size.  This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.  If the training set was not representative, this estimator would produce biased results, and any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.  We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.

>>>>

\section{Methods}
\label{sec:pzcodes}

Here we summarize the twelve \pzpdf\ codes compared in this study, summarized in Table~\ref{tab:list_of_codes}, which include both established and emerging approaches in template fitting and machine learning.
Though not exhaustive, this sample represents codes for which there was sufficient expertise within the \lsstdesc\ Photometric Redshifts Working Group; the authors welcome interest from those outside \lsstdesc\ to have their codes assessed in future investigations that build upon this one.

\begin{table*}  %%% DATA TABLE %%%
\caption{List of \pzpdf\ codes featured in this study} \label{tab:list_of_codes}\resizebox{\textwidth}{!}{
\begin{tabular}{lll}
\hline
\bf Published code & \bf Type & \bf Public source code \\
\hline
\bpz~\citep{Benitez:00} 		   & template fitting	& \url{http://www.stsci.edu/~dcoe/BPZ/} \\
\eazy~\citep{Brammer:08}		   & template fitting & \url{https://github.com/gbrammer/eazy-photoz} \\
\lephare~\citep{Arnouts:99}	   & template fitting	& \url{http://www.cfht.hawaii.edu/~arnouts/lephare.html} \\
\annz~\citep{Sadeh:16}		     & machine learning	& \url{https://github.com/IftachSadeh/ANNZ} \\
\delight~\citep{Leistedt:17}   & machine learning & \url{https://github.com/ixkael/Delight} \\
\flexzboost~\citep{Izbicki:17} & machine learning & \url{https://github.com/tpospisi/flexcode}; \url{https://github.com/rizbicki/FlexCoDE}\\
%\textsc{Frankenz} 	& ML 	& \citet{Speagle:frankenz}	& \url{https://github.com/joshspeagle/frankenz} \\
\gpz~\citep{Almosallam:15b}	   & machine learning	& \url{https://github.com/OxfordML/GPz} \\
\metaphor~\citep{Cavuoti:17}   & machine learning	& \url{http://dame.dsf.unina.it}\\
\cmnn~\citep{Graham:17}        & machine learning & N/A \\
\skynet~\citep{Graff:14}       & machine learning & \url{http://ccpforge.cse.rl.ac.uk/gf/project/skynet/} \\
\tpz~\citep{Carrasco_Kind:13}	 & machine learning	& \url{https://github.com/mgckind/MLZ} \\
\hline
\trainz 	                             & machine learning	& See Section~\ref{sec:method:trainz} \\
\end{tabular}}
\end{table*}

We describe the algorithms and implementations of the model-based and data-driven codes in Sections~\ref{sec:templatecodes} and \ref{sec:trainingcodes} respectively.
For each approach, we note how it accounts for photometric uncertainties, how it interprets negative fluxes, what its native output format of \pzpdf s is, and what, if any, preprocessing, including validation, of the prior information it performs.

\subsection{Template-based Approaches}
\label{sec:templatecodes}

We test three publicly available and commonly used template-based codes, \bpz, \eazy, and \lephare, that share the standard physically motivated approach of calculating model fluxes for a set of template SEDs on a grid of redshift values and evaluating a merit function using the observed and model fluxes.
These three also share the $\chi^{2}$ merit function
\begin{equation}
	\label{eq_temp_chi}
	\chi^{2}(z, T, A) \equiv \sum_{b}^{N_{\rm{filt}}} \left(\frac{F^{\rm{obs}}_{b} - A F^{\rm{mod}}_{b}(T, z)}{\sigma^{\rm{obs}}_{b}}\right)^{2},
\end{equation}
of normalization factor $A$, template $T$, and redshift $z$, which is a sum of observed flux error $\sigma_{b}^{\rm{obs}}$-weighted squared differences between the observed flux $F^{\rm{obs}}_{b}$ and the normalized predicted flux $F^{\rm{mod}}_{b}(T, z)$ in $N_{\rm{filt}}$ photometric filters $b$, which in our case is the \lsst\ $ugrizy$ filters.

@@

\bpz\footnote{\url{http://www.stsci.edu/~dcoe/BPZ/}} \citep[Bayesian Photometric Redshift,][]{Benitez:00} is a template-based \pz\ code that determines the likelihood $p(C \mid z, T)$ of a galaxy's observed colours $C$ for a set of SED templates $T$ at redshifts $z$.
The \bpz\ likelihood is related to Equation~\ref{eq_temp_chi} by $p(C \mid z, T) \propto \exp[- \chi^{2} / 2]$.
Given a Bayesian prior $p(z, T \mid m_{0})$ over apparent magnitude $m_0$ and assuming that the SED templates are spanning and exclusive, \bpz\ constructs the redshift posterior $p(z \mid C, m_0)$ by marginalizing over all SED templates as in \citep[Eq.~3 from][]{Benitez:00}, corresponding to setting the parameter \texttt{PROBS\_LITE=TRUE} in the \bpz\ parameter file.
The \bpz\ prior is the product of an SED template proportion that varies with apparent magnitude $p(T \mid m_{0})$ and a prior $p(z \mid T, m_{0})$ over the expected redshift as a function of apparent magnitude and SED template.

Here we test \bpz-v 1.99.3 with the DC1 template set of Section~\ref{sec:buzztemplates}.
To keep the number of free parameters manageable, the DC1 template set is pre-sorted by the rest-frame $u$-$g$ colour and split into three broad classes of SED template, equivalent to the E, Sp and Im/SB types in .
The Bayesian prior term $p(T \mid m_{0})$ was derived directly from the DC1 training set, and the other term $p(z \mid T, m_{0})$ was chosen to be the best fit for the eleven free parameters of the functional form of \citet{Benitez:00}.
%For photo-$z$ point estimates we use the \texttt{Z\_B} output parameter.
Prior to running the code, the non-detection placeholder magnitude was replaced with an estimate of the one-$\sigma$ detection limit for the undetected band as a proxy for a value close to the estimated sky noise threshold.

@@

In this paper we use the all-templates mode, which fits the photometric data with a linear combination of the five basis templates.
The EAZY architecture cannot accept a template set other than the same five basis templates employed by \texttt{k-correct} when constructing our mock galaxies.
EAZY was thus run using linear combinations of these five templates rather than using the 100 discrete templates shared by the other template-fitting codes.
We employed the $5$ basis templates described in Section~\ref{sec:buzzard}, and set the template error to zero since these same templates were used to produce the simulated catalog photometry. %\red{The quality parameter \texttt{Qz} is...}
The ability to fit for linear combinations of templates highlights an important implementation difference between similar photo-z codes.
The likelihoods include the application of a type-independent apparent magnitude prior estimated from the training data.

@@

\subsection{Training-based Codes}
\label{sec:trainingcodes}

The training-based codes use a variety of algorithms in order to estimate $p(z)$, specifics of each implementation are described in the subsections.
Some aspects of data treatment were left to the individual code runners, for example, whether/how to split the available data with known redshifts into separate training and validation sets.
Another key difference is the treatment of non-detections in one or more bands.
Some codes choose to ignore a band, others replace the value with either an estimate for the detection limit, the mean of other values in the training set, or another default value.
There are varying conventions among training-based codes for treatment of non-detections, and no one prescription dominates in the photo-$z$ literature.
The specific choices for each code affect the results, and contribute to the implicit prior influencing their output.
However, we remind the reader that only 2.0 per cent of our sample has non-detections, almost exclusively in the u-band, and thus should not dominate the code performance differences.

@@

\begin{equation}
p(z|\mathbf{\hat{F}}) \approx \sum_i p(\mathbf{\hat{F}}|z,t_i)\, p(z|t_i)p(t_i),
\end{equation}
\noindent where $p(z|t_i)p(t_i)$ captures prior information about the redshift distributions and abundances of the galaxies, with $t_i$ denoting the galaxy template; while $p(\mathbf{\hat{F}}|z,t_i)$ is the posterior of noisy flux $\mathbf{\hat{F}}$ at redshift $z$.
For each training-target pair, $p(\mathbf{\hat{F}}|z,t_i)$ is evaluated as follows:
\begin{equation} \label{eq:delight_noisy}
p(\mathbf{\hat{F}}|z,t_i) = \int p(\mathbf{\hat{F}}|\mathbf{F})\, p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)\, d\mathbf{F},
\end{equation}
where $p(\mathbf{\hat{F}}|\mathbf{F})$ is the likelihood function, it compares the noisy real flux $\mathbf{\hat{F}}$ with the noiseless flux $\mathbf{F}$ obtained from the linear combination of template models, carefully constructed to account for model uncertainties and different normalization of the same SED; while $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ is the prediction of flux at a different redshift $z$ with respect to the training object with redshift $z_i$ and flux $\mathbf{\hat{F}}_i$. Eq.~\ref{eq:delight_noisy} is essentially the probability that the training and the target galaxies having the same SED but at a different redshift.
The flux prediction $p(\mathbf{F}|z,z_i,\mathbf{\hat{F}}_i)$ of the training galaxy at redshift $z$ is modeled via a Gaussian process,

@@

\textsc{FlexZBoost}\footnote{\url{https://github.com/tpospisi/flexcode};\\ \url{https://github.com/rizbicki/FlexCoDE}} \citep{Izbicki:17} is a particular realization of \texttt{FlexCode}, which is a general-purpose methodology for converting any conditional mean point estimator of $z$ to a conditional {\em density} estimator $f(z \vert \x)$, where $\x$ here represents our photometric covariates and errors.\footnote{Instead of $p(z)$, we use the notation $f(z \vert \x)$ to explicitly show the dependence on $\x$.}
The key idea is to expand the unknown function $f(z \vert \x)$ in an orthonormal basis $\{\phi_i(z)\}_{i}$:
\begin{equation}
f(z|\x)=\sum_{i}\beta_{i }(\x)\phi_i(z). \label{eq::series_expansion}
\end{equation}
By the orthogonality property, the expansion coefficients are just conditional means
\begin{equation}
\beta_{i }(\x) =  \mathbb{E}\left[\phi_i(z)|\x\right] \equiv \int f(z|\x)   \phi_i(z) dz. \label{eq:FlexZBoost_coeffs}
\end{equation}
These coefficients can easily be estimated from data by regression since $\mathbb{E}\left[\phi_i(z)|\x\right]$ is the regression of $\phi_i(z)$ (a transformation of $Z$) on $X$.

@@

The data is prepared for \textsc{GPz} by taking the log of the magnitude errors, decorrelating the data set using PCA and imputing any missing magnitude values using a simple linear model that estimates the missing magnitudes given the observed ones.
The log transformation helps to smooth the long tail distribution of the magnitude errors, which is more stable numerically and makes the optimization process unconstrained.
The missing values are imputed by computing the mean of the training set $\mu$ and its covariance $\Sigma$, then we use the following equation to estimate the missing values from the observed ones %\[
\begin{equation}
x_{u} = \mu_{u}+\Sigma_{uo}\Sigma_{oo}^{-1}(x_{o}-\mu_{o}),
\end{equation}  %\]
where the subscript $o$ in $x_{o}$ indexes the \emph{observed} part of the input $x$, whereas the subscript $u$ indexes the \emph{unobserved} set (similarly for $\mu$ and $\Sigma$).
This is the optimal expected value of the unobserved variables given the observed ones if the distribution is jointly Gaussian, note that if the variables are independent, i.e. $\Sigma_{uo}=0$, this will reduce to a simple average predictor.
We use the Variable Covariance (VC) option in \textsc{GPz} with 200 basis functions after we note that there is no significant increase in the performance on the validation set (using 80\%-20\% training-validation split) and with no cost-sensitive learning applied.

@@

\textsc{METAPhoR} includes data modules for pre-processing, photo-$z$ estimation, and PDF estimation, and post-processing.  The pre-processing includes a model for perturbation of the photometry that is employed in calculating the PDF of the photo-$z$ estimation errors.
The photometric perturbation is defined as:
\begin{equation}
 m_{ij} = m_{ij} + \alpha_{i}F_{ij}u_{\mu=0,\sigma=1}
\end{equation}
where $\alpha_{i}$ is a user selected multiplicative constant (useful in case of multi-survey photometry), $u_{\mu=0,\sigma=1}$ is a random value from the standard normal distribution and $F_{ij}$ is a bimodal function (a constant function + polynomial fitting of the mean magnitude errors on the binned bands), heuristically tuned in such a way that the constant component is the threshold under which the polynomial function is considered too low to provide a significant noise contribution to the photometry perturbation.

@@

\subsection{Simple Ensemble Estimator}\label{sec:method:trainz}

In addition to the main photo-$z$ algorithms described above we also include a very simple method as a pathological example.
For \trainz, as we will we call this simple estimator, we well define $p(z)$ as simply:
\begin{equation}
p(z) = \frac{1}{N_{ \mathrm train}}\sum_{\mathrm i=1}^{N_{\mathrm train}}z_{\mathrm train}
\end{equation}
That is, we simply set the redshift PDF of every galaxy equal to the normalized $N(z)$ of the training sample.
This estimator is essentially a k nearest-neighbour estimator with k equal to the number of galaxies in the training sample.
As the training sample is drawn from the same underlying distribution as the test sample, modulo small deviations due to sample size, the quantiles of the training and test distributions should be identical, modulo fluctuations due to finite sample size.
This is a wildly unrealistic estimator, as it assigns all galaxies, no matter their apparent magnitude, colour, or true redshift, the same redshift PDF, and is thus uninformative at the level of individual object redshifts, but is designed to perform very well for the ensemble of all objects.
If the training set was not representative, this estimator would produce biased results, and any attempts to break up the sample into tomographic bins will fail, as every galaxy has an identical $p(z)$.
We will discuss this method and cautions relative to metrics in Section~\ref{sec:caution}.
